### 2022-03-23 paper1.tex
Final committee edits

% TODO: should I add in some figures to outline these relationships?

@@

Computational social science and disciplines such as sociology and
political science are quick to 
adopt new sources of digital trace data for use in academic research. 
In this article, I examine the validity of Google Search Trends 
as potential indicators of three different cases, namely attitudes, 
disease prevalence and political preferences using eight %TODO double check number
different validated data sources. I use Pearson and Repeated Measures Correlations 
between the Google Trends and the validated indicators as well as multiple linear regression
(for cross-sectional datasets) and random intercept hierarchical linear models
(for longitudinal datasets) as additional tests of the data. 
I fail to find correlation among any of the Google 
Trends tested and their validated indicators. While some Google Trends 
tested were significantly associated with the outcome in the regression models,
effects tended to be small and the total model interpretability remained low,
even when controlling for demographic variables. This article shows that there is no 
external validity of Google Trends for these uses and social scientists
will find no replacement for high quality survey data with Google Trends.

@@

% TODO define big data sources , are we just talking about digital trace data? what is that? Define, even if as footnote. Are 10K pages of field notes big data? Admin data with millions of observations but no text?

% TODO Define digital trace data. I  assume remnants of digital activity archived by corporate or state entities.

New big data sources have led to vast possibilities for social science
research because they are bigger, cheaper, and already available
\citep{kingEnsuringDataRichFuture2011,lazerComputationalSocialScience2009,salganikBitBitSocial2017}.
Before overenthusiastically embracing these
sources and integrating them into our big data toolkits, social scientists must clearly establish
parameters under which these data sources are appropriate, 
and for which questions they are useful 
\citep{bailCulturalEnvironmentMeasuring2014, lazerParableGoogleFlu2014}. As
prior research outlined, the "quantity of data does not mean that one
can ignore foundational issues of measurement and construct validity and
reliability and dependencies among data" 
\citep[p. 1203]{lazerParableGoogleFlu2014}.

With the expansion of big data, some computational social science research has
developed extremely innovative methods that have led to groundbreaking results using 
reliable sources of big data. As an example,\citet{blumenstockPredictingPovertyWealth2015}
use county-level cell-phone records to construct the distribution of poverty
and wealth in Rwanda, a country where national surveys and censuses are
rare and costly. However, \citet{blumenstockPredictingPovertyWealth2015} go to
great lengths to demonstrate that their operationalization of the cell
phone data creates a reliable and valid construct; few social science
papers utilizing big data dig into the construct validity of their
metrics to this extent and even fewer publications focus on
methodological guidelines of how to use sources of big data
\citep[For exceptions, see ][]{asseoTrackingCOVID19Using2020, stilesAssessingCriterionValidity2018}.
However, research has shown the small adjustments to an algorithm or
metric may void any research insight we can pull from such data
\citep{lazerParableGoogleFlu2014}.

@@

Social scientists have yet to widely utilize Google Search Trends as a source of big data.
Google Trends is an open data resource
% TODO Does this just mean free and publicly available?

provided by Google, a major internet search
engine worldwide, that analyzes the top search queries across different localities, 
languages, and time. While Google Trends is freely available and holds a breadth of 
online activities that could potentially be of use in social science, few social 
scientists outside of economics have used this resource \citep[see][for examples]{choi2012predicting, jun2018ten,da2011search}. This paper provides evidence-based guidelines for social scientist who would use these resources in their toolboxes on how we can operationalize 
% TODO: operationalize debated
Google Search Trends
as validated indicators and under what circumstances.

Building on prior research outlining the various issues with different sources of big data
\citep{boydCriticalQuestionsBig2012,lazerIssuesConstructValidity2015},
this paper evaluates the construct validity of Google Search Trends as an indicator of three different
cases, namely attitudes, disease prevalence and political preferences.
These three cases will be tested using attitudinal indicators from the %NORC General Social Survey, 
CDC \citeyearpar{vaches_data} and New York Times \citeyearpar{mask_data}, 
health indicators from the CDC \citeyearpar{suic_data}, 
US cases rates of COVID-19 from The New York Times \citeyearpar{covid_data},  
and political indicators from the historical US Presidential Election results \citeyearpar{pres_data}, 
and the American National Election Survey \citeyearpar{anes_data}. This paper will contribute guidance for creating standards of how to use Google
Trends for social research, and identify potential pitfalls that can compromise the 
analysis including 1. a lack of ecological validity, 2: poor 
% TODO finish this sentence, This is the framework I would use for identifying ‘best practices.’ I would personally just say here is how this can go wrong, here is how you avoid it, and next steps. I could see doing a critical spin off for a place like big data and society.

@@

I will use three categorizations of ways I propose Google Trends could
be operationalized for social scientific usage. First, I test Google
Trends as an operationalization of attitudinal indicators. 
After \citet{bailCulturalEnvironmentMeasuring2014}'s call for
cultural sociologists to utilize the ever-expanding world of big data,
Google trends began slowly appearing as a data source in sociological and
social science research. From research on mass shootings and firearms \citep{brownsteinInternetSearchPatterns2020, semenzaInformationseekingWakeTragedy2020}, 
protest and anti-Muslim sentiment \citep{bailUsingInternetSearch2018,barrieSearchingRacismGeorge2020,grossThereFergusonEffect2017},
to analyzing country-level changes in social perception \citep{reyes_etal18},
Google search trends are a new and innovative indicator of cultural interest.
Extending into social networks and culture, \citet{bailPrestigeProximityPrejudice2019}
even used Google trends to measure how 'culture' spreads around the globe.
% TODO I’ll just say, while I think there is utility here, it is a very particular aspect of culture in a specific populations. I think we are still in the era where the bigness and newness can overshadow truth value.
\subsection{Google Trends and research on health behaviors}

Google Search Trends have been utilized to estimate disease prevalence
and population health in journals like the Journal of
Medical Internet Research. While much of this research has focused on
% TODO I think this may erode your point about increasing legitimacy. If you said JAMA or NEJM, that would be a stronger example. If not, I’d just point to some specific projects then say new journals have emerged but that does not indicate mainstream acceptance…
the COVID-19 pandemic \citep{jimenez_etal20, jimenezCOVID19SymptomGoogle2020, limEstimatingInformationSeekingBehaviour2020, mavraganiCOVID19PredictabilityUnited2020, nguyenGoogleTrendsAnalysis2020, todorovaInternetBasedData2021, mingUnderstandingHealthCommunication2021},
other research has investigated Google Trends as indicators of wellbeing
\citep{brodeurCOVID19LockdownsWellbeing2021, carpiTwitterSubjectiveWellBeing2020, duCOVID19IncreasesOnline2020},
suicidality \citep{burnettTimeTrendsPublic2020}, 
vaccination uptake \citep{dalumhansenEnsembleLearnedVaccination2016}, 
obesity \citep{sarigulNowcastingObesityUsing2014}, 
and even insomnia \citep{zittingGoogleTrendsReveal2020},
to cover a few examples. For a partial review of other utilizations, see \citet{nutiUseGoogleTrends2014}. According to \citet{jaidkaInformationseekingVsSharing2021}, the majority of studies profess
a correlation of \> .70 between the IV and DV, "demonstrating the vast potential of Google
Search as a proxy for monitoring population health" (p. 3) based on
assumptions that individuals search because of self-diagnosis and to
identify possible courses of treatment \citep{dechoudhurySeekingSharingHealth2014}.

@@

Various sources have also used Google Trends to forecast
political elections and political attitudes
\citep{wolfTrendingRightDirection2018}. For instance,
\citet{swearingenGoogleInsightsSenate2014} investigate how U.S. Senate
Elections relate to attention measured by search traffic.
\citet{prado-romanGoogleTrendsPredictor2020} demonstrate how Google Search trends
successfully predicts presidential election results in both the United
States and Canada. Finally, the OECD Development Centre is investigating
how Google data can help elucidate governments' approval in Latin
America \citep{montoyaUsingGoogleData2020}. 
% TODO what are the results of these papers? do they show promise?

@@

\citet{cebrian_domenech22} emphasize in their research that there are some 
non-negligible issues with the measurement of Google Trends, pointing to 
errors within the purview of accuracy, completeness, consistency, and validity
% TODO I know it is basic stats, but I would provide a footnote explaining each of your terms. Here, these line up with mainline stats but disciplines vary.

\citep{KarrDataQuality}. \citet{cebrian_domenech22} point not only to 
meta issues like the lack of transparency into the sampling methods 
and population coverage, but they identify real differences 
in the data provided by the Google Trends API for the same 
researcher-entered query, pointing 
to likely internal process used by Google to sample their data to compute the trends.
These authors show that the variability of the results over time for the same query
can alter the results of analysis and forecasting significantly; they propose
caution in the use of the digital trace data for other researchers until
further research can determine the cause for the inaccuracies. 
% TODO Yes, big issue. But, it is also seen in machine learning, survey administration, etc. I’d assume one solution for dealing with fiddly output is to just sample repetedly and do some sort of average, perhaps with a correction if the direction of the error is non random.

Finally, it is important to note that most of the research using 
Google Trends as digital trace data do so to indicate
population's \textit{attention} to a search query. Because
% TODO I’d also say, the population is examined differently. It is really a population of action remnants being used to get at a population of people, which is presumed to be represented online.
searches can be done with many different intentions, such as 
information gathering, navigational purposes, etc., it may
be hard to declare that an overall level of engagement with a topic
on Google Search may be indicative of anything more than interest \citep{da2011search}.
To reference Clifford Geertz's invocation of Ryle, we only know that the search was 
performed, not whether it indicated a digital wink or twitch \citeyearpar{geertz1973thick}.
%TODO CITE above, For a soc audience some sort of reference to the dilemma of interpretation would be interesting.

@@

%TODO There are a lot of validities here. Be precise and explain what you are concretely examining. define which validity here 
To investigate the construct and criterion validity of the use of Google Trends
in these three areas, I gathered geo-located social science data across multiple
sources to address the three areas of inquiry for this paper.  Table
\ref{tab:data-sources-table} outlines which data sources I use in this
project and which trends are matched to each source.

@@

% TODO create a  table People will want to see a table of variables, their prompts, and ranges.

@@

I also use U.S. COVID-19 rates to validate health and disease related topics. I
retrieve U.S. county-level COVID-19 rates from \citet{covid_data}, who compile this
data based on reports from state and local health agencies. It is widely
acknowledged that there are biases in this data due to inconsistencies and
availability in testing as well as different community propensity to test
% TODO Also blatant political interference in some places, which is probably worth acknowledging.
\citep{gu22, cdc20a} However, it is the best measure we have of actual case rates
without extensive adjustments.
Case Rate is measured as number of cases per 100,000 population. Observations
vary from a minimum of zero to a maximum of 1460.46 for each Monday from January 27, 2020, through
December 27, 2021. There are 285,986 cases across 3,136 counties and 101
dates. Missing data were interpolated using The Zoo package for R \citep{zoo}.

@@

Finally, I evaluate Google Trends as an operationalization of political
attitudes by first looking at actual voting outcomes in historical US
Presidential Election results. This data comes from  \citet{pres_data}, who scraped the
results from Townhall.com, Fox News, Politico, and the New York Times. Data on 
presidential outcomes were available for 3,147 counties in 2016 and
3,118 counties in 2020. Each variable measures the percent of votes
for the candidate, with the lowest percent at 3.09\% and the highest at 92.15\%.

In addition, I use data on political opinions from the American National
Election Survey 2020 Time-Series Study \citep{anes_data}. This data is paired with
restricted data provided by ICPSR, the Inter-university Consortium for Political
and Social Research, which contains the geo-ids of the 5,441
survey respondents. The study interviewed respondents in a pre-election survey
that was conducted between August 18, 2020, and the day of the US Presidential
election, November 3, 2020.  %TODO
I will write about actual variables when my data extract request is approved.

% TODO: WHICH VARIABLES
% TODO: VARIABLE SCALE

@@

In addition to specific measures of interest, I also gathered vadata on potentially
confounding variables. The first Seven of these
variables come from the 2010-2019 5-year American Community Surveys
\citep{acs2019, acs2018, acs2017, acs2016, acs2015, acs2014, acs2013, acs2012, acs2011, acs2010}. 
These data include total population, population density,
unemployment rate for those sixteen years old and over, median county income, average commute
time, percent of households living under the poverty line, and percent above 65
years old. In addition, some models employ the U.S. Current Population Survey \&
American Community Survey Geographic Estimates of Internet Use, 1997-2018
\citep{internet_use} to estimate households with broadband internet subscriptions.
This final variable is an attempt to capture the latent propensity to use the
internet for information search.

@@

To address longitudinal correlations, I employ Repeated Measures Correlation
using the \texttt{rmcorr} package in \texttt{r} \citep{bland1995, bakdash2017}. 
Repeated Measures Correlation is useful for determining the within-county
association for paired measures across time and counties. Results of these 
correlations are in table \ref{tab:corr-results}.

As an additional criteria to evaluate the strength of the relationships, 
I employ multiple linear regression
(for cross-sectional datasets) and random intercept hierarchical linear models
\citep{pinheiro_etal21} (for longitudinal datasets) to identify the strength of
relationships across locales.
%TODO This all sounds fine to me, but I’m less up to date on model specification and nesting, so I’m not the best to ask. I would also specify why these models are appropriate in one sentence max because methods people about these things. Also,they may have sampling assumptions the data violate so clarifying that that is not an issue may be necessary.   
I include location metadata like county population size,
broadband rates, and median income to adjust for potentially
confounded relationships between Google search Trends and credible
indicators. Identifying these possible confounded relationships may help to
explain why some articles find relationships between the trends and outcomes
while others did not.
% TODO Why? To have a standardized net impact? Sociologists usually have a strong negative reaction to this. In contrast, in psych they do this with everything.  
For linear regression models, I normalize independent
variables; in other words, variables have been centered and scaled to have a mean of zero and
standard deviation of one. Results of these models are in tables \ref{tab:vacc_hes_analysis} through \ref{tab:pres_2020_analysis}.

@@

The first attitudinal indicator is Vaccine Hesitancy for COVID-19 vaccines
\citep{vaches_data}. When comparing the measure of vaccine hesitancy to four
different Google Search Trends, the correlation does not exceed -0.4657
('Coronavirus (Disease)' and vaccine hesitancy). Correlation's under |0.40| are 
weak according to the common standards. With only one correlation 
above that threshold, I conclude that the overall trends are at best weakly
correlated with these attitudes. When running these
correlations in multiple linear regression (see the results in table
\ref{tab:vacc_hes_analysis}), I see an $r^2$ of 0.233 (Model 1),
indicating that the Google Trends explains about 23\% of
the variation in Vaccine Hesitancy alone. Demographic characteristics like the
percentage of households with broadband internet and the population density are
able to explain about 35\% of the variation (model 2), outperforming the first
models. The Trend coefficients for the trends themselves, however, are
significant and remain significant when controlling for demographics. This reinforces the finding from the Pearson Correlation that there is a significant but weak relationship between the Google Trends and Vaccine Hesitancy.

@@

\input{tables/paper1/vacc_hes_analysis}

@@

\input{tables/paper1/mask_analysis}

@@

In addition to attitudinal measures, I attempt to validate Google Trends for
uses in the measurements of health indicators. The first indicator I assess is
COVID-19 case rates from 2020 through 2021. The Pearson correlation results
indicate negligible to weak relationships between the search terms and the
actual case rates across time and place. The repeated measures correlation
coefficient, a more reliable and less biased measure for longitudinal
correlations, reveals similarly weak results with a maximum correlation of 0.32
between 'smell loss' and rates of COVID-19. To further investigate the relationship, I
include the three trends in random intercept hierarchical linear models. Model 1
of table \ref{tab:covid_analysis} indicates that the trends themselves are able
to explain about 20\% of the variation in COVID-19 rates ($r^2$ = 0.201). Few
of the demographic variables influence COVID-19 rates in models 2 or 3,
though there is some suppression for median income, unemployment rates,
and population density. Model 3 has an $r^2$ of 0.200. I conclude that the
Google Trends are slightly related to COVID-19 rates, but that this relationship
is weak, and we should not attempt to use Google Trends as an indicator of actual
health outcomes.

\input{tables/paper1/covid_analysis}

Another health outcome I test is county suicide rates. As with previous cases,
there is only negligible correlation between actual suicide rates and Google
Search Trends, not exceeding 0.0845. In the more accurate repeated measures
correlation coefficient we see a maximum of 0.142 correlation, remaining
negligible. I introduce these trends into a random intercept model to control
for inter-group variation over time in table \ref{tab:suicide_analysis}. While
table \ref{tab:suicide_analysis} model 1 reveals a statistically significant relationship
between suicide hotline trends and suicide rates, the effect is small, and the
entire model only has an $r^2$ of 0.058. Adding in demographic features in model
2 improves the $r^2$ to 0.56. In model 3, the inclusion of Google Trends
decreases the overall amount of suicide rate variance explained
compared to model 2. These analyses lead me to conclude that Google Search
Trends should not be used by researchers as indicators of suicide rates or intentions.

\input{tables/paper1/suicide_analysis}

@@

In table \ref{tab:pres_2016_analysis}, I outline the results for the model
predicting the county percentage of votes for Hillary Clinton. Model 1
demonstrates how well Google Trends predict the votes; while the
trend for 'Hillary Clinton' is significantly associated with votes, the model's
$r^2$ ($r^2$ = 0.027) shows that it does little to help explain the model
variance. On the other hand, adding in demographic features improves the model
fit quite well, bringing the $r^2$ up to 0.313 in model 2 and $r^2$ 0.327 in
model 3. Table \ref{tab:pres_2020_analysis} shows this same analysis for the
2020 election and predicts the percentage of votes for Joe Biden. While this
model is analogous to those in table \ref{tab:pres_2016_analysis}, this
model 1 demonstrates how unrelated Google Trends can be from actual outcomes. In
conclusion, I find that Google Trends are an unreliable and invalid indicator of
political support.

>>>>

@@

Computational social science and disciplines such as sociology and
political science are often quick to adopt new sources of digital 
trace data for use in academic research. 
In this article, I examine the validity of Google Search Trends 
as potential indicators of three different cases, namely attitudes, 
disease prevalence and political preferences using five
different validated data sources. I use Pearson and Repeated Measures Correlations 
between the Google Trends and the validated indicators as well as multiple linear regression
(for cross-sectional datasets) and random intercept hierarchical linear models
(for longitudinal datasets) as additional tests of the data. 
I fail to find correlation among any of the Google 
Trends tested and their validated indicators. While some Google Trends 
tested were significantly associated with the outcome in the regression models,
effects tended to be small and the total model interpretability remained low,
even when controlling for demographic variables. This article shows that there is no 
external validity of Google Trends for these uses and social scientists
will find no replacement for high quality survey data with Google Trends.

@@

New sources of digital trace data have led to vast possibilities for social science
research because they are bigger, cheaper, and already available 
\citep{kingEnsuringDataRichFuture2011,lazerComputationalSocialScience2009,salganikBitBitSocial2017}.
Digital trace data are the expansive volumes of information that are now available with the
introduction and widespread adoption of the internet and include sources such as  
social media sites, search data, blogs, adminstrative data on websites, the internet 
archives, etc. In short, digital trace data are the crumbs that are left as a person
peruses and interacts in digital spaces. Scholars have noted that these data provide
many new research avenues for social scientists because they are always on, non-reactive to
a researcher's observation, and very often captual social interactions and relationships,
but warn they can be Inaccessible, Non-Representative, and be prone to drifting due to
algorithmic confounding \citep{salganikBitBitSocial2017}.
Before overenthusiastically embracing these
sources and integrating them into our big data toolkits, social scientists must clearly establish
parameters under which these data sources are appropriate, 
and for which questions they are useful 
\citep{bailCulturalEnvironmentMeasuring2014, lazerParableGoogleFlu2014}. As
prior research outlined, the "quantity of data does not mean that one
can ignore foundational issues of measurement and construct validity and
reliability and dependencies among data" 
\citep[p. 1203]{lazerParableGoogleFlu2014}.

With the expansion of big data, some computational social science research has
developed extremely innovative methods that have led to groundbreaking results using 
reliable sources of big data. As an example,\citet{blumenstockPredictingPovertyWealth2015}
use county-level cell-phone records to construct the distribution of poverty
and wealth in Rwanda, a country where national surveys and censuses are
rare and costly. However, \citet{blumenstockPredictingPovertyWealth2015} go to
great lengths to demonstrate that their use of the cell
phone data creates a reliable and valid construct; few social science
papers utilizing big data dig into the construct validity of their
metrics to this extent and even fewer publications focus on
methodological guidelines of how to use sources of big data
\citep[For exceptions, see ][]{asseoTrackingCOVID19Using2020, stilesAssessingCriterionValidity2018}.
However, research has shown the small adjustments to an algorithm or
metric may void any research insight we can pull from such data
\citep{lazerParableGoogleFlu2014}.

@@

Social scientists have yet to widely utilize Google Search Trends as a source of big data.
Google Trends is a free and publicly accessible data resource
provided by Google, a major internet search
engine worldwide, that analyzes the top search queries across different localities, 
languages, and time. While Google Trends is freely available and holds a breadth of 
online activities that could potentially be of use in social science, few social 
scientists outside of economics have used this resource \citep[see][for examples]{choi2012predicting, jun2018ten,da2011search}. This paper provides evidence-based guidelines for social scientist who would use these resources in their toolboxes on how we can triangulate Google Search Trends
as validated indicators and under what circumstances.

Building on prior research outlining the various issues with different sources of big data
\citep{boydCriticalQuestionsBig2012,lazerIssuesConstructValidity2015},
this paper evaluates the construct validity of Google Search Trends as an indicator of three different
cases, namely attitudes, disease prevalence and political preferences.
These three cases will be tested using attitudinal indicators from the %NORC General Social Survey, 
CDC \citeyearpar{vaches_data} and New York Times \citeyearpar{mask_data}, 
health indicators from the CDC \citeyearpar{suic_data}, 
US cases rates of COVID-19 from The New York Times \citeyearpar{covid_data},  
and political indicators from the historical US Presidential Election results \citeyearpar{pres_data}.
% , and the American National Election Survey \citeyearpar{anes_data}.
This paper will contribute guidance for creating standards of how to use Google
Trends for social research, and identify potential pitfalls that can compromise the 
analysis including the focus of this paper, which shows that Google Trends has a
lack of ecological validity.

@@

I will use three categorizations of ways I propose Google Trends could
be utilized for social scientific usage. First, I test Google
Trends as an indicator of attitudinal indicators. 
After \citet{bailCulturalEnvironmentMeasuring2014}'s call for
cultural sociologists to utilize the ever-expanding world of big data,
Google trends began slowly appearing as a data source in sociological and
social science research. From research on mass shootings and firearms \citep{brownsteinInternetSearchPatterns2020, semenzaInformationseekingWakeTragedy2020}, 
protest and anti-Muslim sentiment \citep{bailUsingInternetSearch2018,barrieSearchingRacismGeorge2020,grossThereFergusonEffect2017},
to analyzing country-level changes in social perception \citep{reyes_etal18},
Google search trends are a new and innovative indicator of cultural interest.
Extending into social networks and culture, \citet{bailPrestigeProximityPrejudice2019}
even used Google trends to measure how they saw 'culture' spreads around the globe
through the proliferation of search terms.

\subsection{Google Trends and research on health behaviors}

Google Search Trends have been utilized to estimate disease prevalence
and population throughout the health sciences. While much of this research has focused on
the COVID-19 pandemic \citep{jimenez_etal20, jimenezCOVID19SymptomGoogle2020, limEstimatingInformationSeekingBehaviour2020, mavraganiCOVID19PredictabilityUnited2020, nguyenGoogleTrendsAnalysis2020, todorovaInternetBasedData2021, mingUnderstandingHealthCommunication2021},
other research has investigated Google Trends as indicators of wellbeing
\citep{brodeurCOVID19LockdownsWellbeing2021, carpiTwitterSubjectiveWellBeing2020, duCOVID19IncreasesOnline2020},
suicidality \citep{burnettTimeTrendsPublic2020}, 
vaccination uptake \citep{dalumhansenEnsembleLearnedVaccination2016}, 
obesity \citep{sarigulNowcastingObesityUsing2014}, 
and even insomnia \citep{zittingGoogleTrendsReveal2020},
to cover a few examples. For a partial review of other utilizations, see \citet{nutiUseGoogleTrends2014}. According to \citet{jaidkaInformationseekingVsSharing2021}, the majority of studies profess
a correlation of \> .70 between the IV and DV, "demonstrating the vast potential of Google
Search as a proxy for monitoring population health" (p. 3) based on
assumptions that individuals search because of self-diagnosis and to
identify possible courses of treatment \citep{dechoudhurySeekingSharingHealth2014}.

@@

Various sources have also used Google Trends to forecast
political elections and political attitudes
\citep{wolfTrendingRightDirection2018}. For instance,
\citet{swearingenGoogleInsightsSenate2014} investigate how U.S. Senate
Elections relate to attention measured by search traffic.
\citet{pradoromanGoogleTrendsPredictor2020} demonstrate how Google Search trends
successfully predicts presidential election results in both the United
States and Canada. Finally, the OECD Development Centre is investigating
how Google data can help elucidate governments' approval in Latin
America \citep{montoyaUsingGoogleData2020}. Overall, the papers 
cited here find favorable relationships between Google Trends and
voting outcomes and claim that it is also a valid measure of public attention.

@@

\citet{cebrian_domenech22} emphasize in their research that there are some 
non-negligible issues with the measurement of Google Trends, pointing to 
errors within the purview of accuracy, completeness, consistency, and validity,
as outlined in \citet{KarrDataQuality}.  
% I know it is basic stats, but I would provide a footnote explaining each of your terms. Here, these line up with mainline stats but disciplines vary.
 \citet{cebrian_domenech22} point not only to 
meta issues like the lack of transparency into the sampling methods 
and population coverage, but they identify real differences 
in the data provided by the Google Trends API for the same 
researcher-entered query, pointing to likely internal process 
used by Google to sample their data to compute the trends.
These authors show that the variability of the results over time for the same query
can alter the results of analysis and forecasting significantly. While the 
issue of unreliability isn't unique to the Google Trends API, 
\citet{cebrian_domenech22} propose 
caution in the use of this digital trace data for other researchers until
further research can determine the root cause for the inaccuracies.
% Yes, big issue. But, it is also seen in machine learning, survey 
% administration, etc. I’d assume one solution for dealing with fiddly 
% output is to just sample repetedly and do some sort of average, 
% perhaps with a correction if the direction of the error is non random.

Finally, it is important to note that most of the research using 
Google Trends as digital trace data do so to indicate
population's \textit{attention} to a search query. Because
% I’d also say, the population is examined differently. It is really a population of action remnants being used to get at a population of people, which is presumed to be represented online.
searches can be done with many different intentions, such as 
information gathering, navigational purposes, etc., it may
be hard to declare that an overall level of engagement with a topic
on Google Search may be indicative of anything more than interest \citep{da2011search}.
To reference Clifford Geertz's invocation of Ryle, we only know that the search was 
performed, not whether it indicated a digital wink or twitch \citeyearpar{geertz1973thick}.
% For a soc audience some sort of reference to the dilemma of interpretation would be interesting.

@@

To test how well Google Trends can be used to indicate other measures, 
I use various tests to validate the criterion validity of the trends. 
Criterion validity measures how well one measure, an operationalization of a 
specific construct, is able to predict a theoretical representation 
of the construct — the criterion. For this article, I employ 
concurrent criterion validity which looks at how Google Trends can
predict various criterions in the same time scale. 
I gathered geo-located social science data across multiple
sources to address the three areas of inquiry for this paper.  Table
\ref{tab:data-sources-table} outlines which data sources I use in this
project and which trends are matched to each source.

@@

%  create a  Table People will want to see a Table of variables, their prompts, and ranges.

@@

I also use U.S. COVID-19 rates to validate health and disease related topics. I
retrieve U.S. county-level COVID-19 rates from \citet{covid_data}, who compile this
data based on reports from state and local health agencies. It is widely
acknowledged that there are biases in this data due to inconsistencies and
availability in testing as well as different community propensity to test
%  Also blatant political interference in some places, which is probably worth acknowledging.
\citep{gu22, cdc20a} However, it is the best measure we have of actual case rates
without extensive adjustments.
Case Rate is measured as number of cases per 100,000 population. Observations
vary from a minimum of zero to a maximum of 1460.46 for each Monday from January 27, 2020, through
December 27, 2021. There are 285,986 cases across 3,136 counties and 101
dates. Missing data were interpolated using The Zoo package for R \citep{zoo}.

@@

Finally, I evaluate Google Trends as an indicator of political
attitudes by first looking at actual voting outcomes in historical US
Presidential Election results. This data comes from  \citet{pres_data}, who scraped the
results from Townhall.com, Fox News, Politico, and the New York Times. Data on 
presidential outcomes were available for 3,147 counties in 2016 and
3,118 counties in 2020. Each variable measures the percent of votes
for the candidate, with the lowest percent at 3.09\% and the highest at 92.15\%.

% In addition, I use data on political opinions from the American National
% Election Survey 2020 Time-Series Study \citep{anes_data}. This data is paired with
% restricted data provided by ICPSR, the Inter-university Consortium for Political
% and Social Research, which contains the geo-ids of the 5,441
% survey respondents. The study interviewed respondents in a pre-election survey
% that was conducted between August 18, 2020, and the day of the US Presidential
% election, November 3, 2020.  %TODO
% I will write about actual variables when my data extract request is approved.

% % TODO: WHICH VARIABLES
% % TODO: VARIABLE SCALE

@@

In addition to specific measures of interest, I also gathered data on potentially
confounding variables. The first Seven of these
variables come from the 2010-2019 5-year American Community Surveys
\citep{acs2019, acs2018, acs2017, acs2016, acs2015, acs2014, acs2013, acs2012, acs2011, acs2010}. 
These data include total population, population density,
unemployment rate for those sixteen years old and over, median county income, average commute
time, percent of households living under the poverty line, and percent above 65
years old. In addition, some models employ the U.S. Current Population Survey \&
American Community Survey Geographic Estimates of Internet Use, 1997-2018
\citep{internet_use} to estimate households with broadband internet subscriptions.
This final variable is an attempt to capture the latent propensity to use the
internet for information search.

@@

To address longitudinal correlations, I employ Repeated Measures Correlation
using the \texttt{rmcorr} package in \texttt{r} \citep{bland1995, bakdash2017}. 
Repeated Measures Correlation is useful for determining the within-county
association for paired measures across time and counties. Results of these 
correlations are in Table \ref{tab:corr-results}.

As an additional criteria to evaluate the strength of the relationships, 
I employ multiple linear regression
(for cross-sectional datasets) and fixed effect hierarchical linear models
\citep{pinheiro_etal21} (for longitudinal datasets) to identify the strength of
relationships across locales. Multiple linear regression is appropriate because 
the cross-sectional data are non-nested which a continous outcome. Hierarchical
linear models are necessary to control for the between-subject variation 
in the longitudinal models. Fixed models are appropriate because 
they extract the between group-differences from the model to allow the reader to 
focus on the overall effects of each variable. 
In these models, include location metadata like county population size,
broadband rates, and median income to adjust for potentially
confounded relationships between Google search Trends and credible
indicators. Identifying these possible confounded relationships may help to
explain why some articles find relationships between the trends and outcomes
while others did not.
For linear regression models, I normalize independent
variables so that each variable across each model is on the same scale and can be compared. 
In other words, variables have been centered and scaled to have a mean of zero and
standard deviation of one. Results of these models are in Tables \ref{tab:vacc_hes_analysis} through \ref{tab:pres_2020_analysis}.

@@

The first attitudinal indicator is Vaccine Hesitancy for COVID-19 vaccines
\citep{vaches_data}. When comparing the measure of vaccine hesitancy to four
different Google Search Trends, the correlation does not exceed -0.4657
('Coronavirus (Disease)' and vaccine hesitancy). Correlation's under |0.40| are 
weak according to the common standards. With only one correlation 
above that threshold, I conclude that the overall trends are at best weakly
correlated with these attitudes. When running these
correlations in multiple linear regression (see the results in Table
\ref{tab:vacc_hes_analysis}), I see an $r^2$ of 0.233 (Model 1),
indicating that the Google Trends explains about 23\% of
the variation in Vaccine Hesitancy alone. Demographic characteristics like the
percentage of households with broadband internet and the population density are
able to explain about 35\% of the variation (model 2), outperforming the first
models. The Trend coefficients for the trends themselves, however, are
significant and remain significant when controlling for demographics. This reinforces the finding from the Pearson Correlation that there is a significant but weak relationship between the Google Trends and Vaccine Hesitancy.

@@

\input{tables/paper1/vacc_hes_analysis}
\begin{figure}[h]
{\centering \includegraphics[width=0.8\linewidth]{figs/paper1/vacc_hes_plot-1.pdf}}
\caption{Vaccine hesitancy by each Google Trend with a fitted linear regression line}\label{fig:vacc_hes_plot}
\end{figure}

@@

\input{tables/paper1/mask_analysis}
\begin{figure}[h]
{\centering \includegraphics[width=0.8\linewidth]{figs/paper1/mask_plot-1.pdf}}
\caption{Rare mask usage by each Google Trend with a fitted linear regression line}\label{fig:mask_plot-1}
\end{figure}

@@

In addition to attitudinal measures, I attempt to validate Google Trends for
uses in the measurements of health indicators. The first indicator I assess is
COVID-19 case rates from 2020 through 2021. The Pearson correlation results
indicate negligible to weak relationships between the search terms and the
actual case rates across time and place. The repeated measures correlation
coefficient, a more reliable and less biased measure for longitudinal
correlations, reveals similarly weak results with a maximum correlation of 0.32
between 'smell loss' and rates of COVID-19. To further investigate the relationship, I
include the three trends in random intercept hierarchical linear models. Model 1
of Table \ref{tab:covid_analysis} indicates that the trends themselves are able
to explain about 20\% of the variation in COVID-19 rates ($r^2$ = 0.201). Few
of the demographic variables influence COVID-19 rates in models 2 or 3,
though there is some suppression for median income, unemployment rates,
and population density. Model 3 has an $r^2$ of 0.200. I conclude that the
Google Trends are slightly related to COVID-19 rates, but that this relationship
is weak, and we should not attempt to use Google Trends as an indicator of actual
health outcomes.

\input{tables/paper1/covid_analysis}
\begin{figure}[h]
{\centering \includegraphics[width=0.8\linewidth]{figs/paper1/covid_plot-1.pdf}}
\caption{Covid case rate for 09-14-2020 by each Google Trend with a fitted linear regression line}\label{fig:covid_plot-1}
\end{figure}

Another health outcome I test is county suicide rates. As with previous cases,
there is only negligible correlation between actual suicide rates and Google
Search Trends, not exceeding 0.0845. In the more accurate repeated measures
correlation coefficient we see a maximum of 0.142 correlation, remaining
negligible. I introduce these trends into a random intercept model to control
for inter-group variation over time in Table \ref{tab:suicide_analysis}. While
table \ref{tab:suicide_analysis} model 1 reveals a statistically significant relationship
between suicide hotline trends and suicide rates, the effect is small, and the
entire model only has an $r^2$ of 0.058. Adding in demographic features in model
2 improves the $r^2$ to 0.56. In model 3, the inclusion of Google Trends
decreases the overall amount of suicide rate variance explained
compared to model 2. These analyses lead me to conclude that Google Search
Trends should not be used by researchers as indicators of suicide rates or intentions.

\input{tables/paper1/suicide_analysis}
\begin{figure}[h]
{\centering \includegraphics[width=0.8\linewidth]{figs/paper1/suicide_plot-1.pdf}}
\caption{Suicide death rate for 2019 by each Google Trend with a fitted linear regression line}\label{fig:suicide_plot-1}
\end{figure}

@@

In Table \ref{tab:pres_2016_analysis}, I outline the results for the model
predicting the county percentage of votes for Hillary Clinton. Model 1
demonstrates how well Google Trends predict the votes; while the
trend for 'Hillary Clinton' is significantly associated with votes, the model's
$r^2$ ($r^2$ = 0.027) shows that it does little to help explain the model
variance. On the other hand, adding in demographic features improves the model
fit quite well, bringing the $r^2$ up to 0.313 in model 2 and $r^2$ 0.327 in
model 3. Table \ref{tab:pres_2020_analysis} shows this same analysis for the
2020 election and predicts the percentage of votes for Joe Biden. While this
model is analogous to those in Table \ref{tab:pres_2016_analysis}, this
model 1 demonstrates how unrelated Google Trends can be from actual outcomes. In
conclusion, I find that Google Trends are an unreliable and invalid indicator of
political support.
